---
title: "Zillow EDA Tax Prices"
output: html_notebook
---

####This is my first kaggle kernel.
####It attempts to follow along from Philipp Spachtholz's Exploratory Analysis - Zillow
####using data.tables when possible instead of dplyr.

####Goal is to ultimately use the tax information + lat lon to do some cool mapping and learn leaftlet



Loads packages
```{r setup, echo=FALSE}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
getwd()
library(ggplot2)
library(DescTools)
library(lubridate)
library(plyr)
library(dplyr)
library(magrittr)
library(ggthemes)
library(data.table)
library(caret)
library(viridis)
library(leaflet)

theme_set(theme_economist_white())
options(scipen = 0)
```
<!--
Loads in data using fread from data.table
```{r ghostdataload, include=FALSE, echo=FALSE}
raw_properties <- as.data.table(fread('../input/properties_2016.csv', showProgress = TRUE))

raw_train <- as.data.table(fread('../input/train_2016.csv', showProgress = TRUE))

#useless dataframe for now, formatting for later
raw_sample_submission <- as.data.table(fread('../input/sample_submission.csv', showProgress = TRUE))

```
-->




<!--
Creates smaller dataframe that won't crash
```{r reducefilesize, include=FALSE, echo=FALSE}



set.seed(1618)
reductionsize <- .10

smallIndex <- createDataPartition(raw_train$parcelid, p = reductionsize,  list = FALSE, times = 1)

small_train<- raw_train[smallIndex]

#need to learn how to better way to subset in data.tables
small_properties <- (raw_properties[parcelid %in% small_train$parcelid])

saveRDS(raw_properties,'../output/properties.rds')
saveRDS(raw_train,'../output/train.rds')
saveRDS(small_properties,'../output/small_properties.rds')
saveRDS(small_train,'../output/small_train.rds')

```
-->


```{r maindataload, echo=FALSE}
#df_properties <- readRDS('../output/raw_properties.rds')
#df_transactions <- readRDS('../output/raw_train.rds')

df_properties <- readRDS('../output/small_properties.rds')
df_transactions <- readRDS('../output/small_train.rds')

```

Makes our data more user friendly by renaming colnames
```{r dataprep_renamecol, echo=FALSE, include=FALSE}


#colname changes copied from Philipp Spachtholz's Exploratory Analysis - Zillow
# converted to data.tables setnames function instead of rename in dplyr
setnames(df_properties, old=c( 'parcelid',
                      'yearbuilt',
                      'basementsqft',
                      'yardbuildingsqft17',
                      'yardbuildingsqft26',
                      'poolsizesum',
                      'lotsizesquarefeet',
                      'garagetotalsqft',
                      'finishedfloor1squarefeet',
                      'calculatedfinishedsquarefeet',
                      'finishedsquarefeet6',
                      'finishedsquarefeet12',
                      'finishedsquarefeet13',
                      'finishedsquarefeet15',
                      'finishedsquarefeet50',
                      'unitcnt',
                      'numberofstories',
                      'roomcnt',
                      'bathroomcnt',
                      'bedroomcnt',
                      'calculatedbathnbr',
                      'fullbathcnt',
                      'threequarterbathnbr',
                      'fireplacecnt',
                      'poolcnt',
                      'garagecarcnt',
                      'regionidcounty',
                      'regionidcity',
                      'regionidzip',
                      'regionidneighborhood',
                      'fireplaceflag',
                      'hashottuborspa',
                      'buildingqualitytypeid',
                      'buildingclasstypeid',
                      'typeconstructiontypeid',
                      'decktypeid',
                      'storytypeid',
                      'heatingorsystemtypeid',
                      'airconditioningtypeid',
                      'architecturalstyletypeid',
                      'taxvaluedollarcnt',
                      'landtaxvaluedollarcnt',
                      'structuretaxvaluedollarcnt',
                      'assessmentyear'), 
         
              new=c('id_parcel',
                    'year_built',
                    'area_basement',
                    'area_patio',
                    'area_shed',
                    'area_pool',
                    'area_lot',
                    'area_garage',
                    'area_firstfloor_finished',
                    'area_total_calc',
                    'area_base',
                    'area_live_finished',
                    'area_liveperi_finished',
                    'area_total_finished',
                    'area_unknown',
                    'num_unit',
                    'num_story',
                    'num_room',
                    'num_bathroom',
                    'num_bedroom',
                    'num_bathroom_calc',
                    'num_bath',
                    'num_75_bath',
                    'num_fireplace',
                    'num_pool',
                    'num_garage',
                    'region_county',
                    'region_city',
                    'region_zip',
                    'region_neighbor',
                    'flag_fireplace',
                    'flag_tub',
                    'quality',
                    'framing',
                    'material',
                    'deck',
                    'story',
                    'heating',
                    'aircon',
                    'architectural_style',
                    'tax_totalvalue',
                    'tax_landvalue',
                    'tax_structurevalue',
                    'year_taxasessment'
              )
         )

#converts df_transactions
setnames(df_transactions, old = c('parcelid', 'transactiondate'),
                          new = c('id_parcel', 'date_of_transact')
         )

```

###Explores transactions by timeframe.

Pretty boring so far, no real insights. 
```{r edatransaction}
df_transactions[, date_of_transact:= as.Date(date_of_transact)]
df_transactions[, month_of_transact:= month(date_of_transact)]
df_transactions[, year_of_transact:= year(date_of_transact)]
df_transactions[, year_month_transact:= make_date(year=year(date_of_transact),month=month(date_of_transact))]

#need to add unique() here
#transacts_per_year <- df_transactions[, N := uniqueN(id_parcel), by = year_month_transact]

transacts_per_year <- df_transactions[, .N, by = year_month_transact]
transacts_per_day <- df_transactions[, .N, by = date_of_transact]


ggplot(transacts_per_year,aes(x=year_month_transact,y=N)) + geom_bar(stat="identity", fill="dodgerblue2")

ggplot(transacts_per_day,aes(x=date_of_transact,y=N)) + geom_bar(stat="identity", fill="dodgerblue2")


```

Plots some exploratory analysis by tax data

First up a frequency plot by the different tax values for land and structure
```{r taxpricehistogram, fig.width=12, fig.height=8}
long_tax <- melt(df_properties,
                 id.vars = c("id_parcel"),
                 measure.vars = c("tax_totalvalue", "tax_landvalue", "tax_structurevalue"),
                 variable.name = "tax_type", 
                 value.name = "tax_value")

str(long_tax)

col_pal <- "Set1"

ggplot(long_tax, aes(x=tax_value, colour = tax_type)) +
  geom_line(stat="density", size=1.2) +   
  theme_bw()+
  scale_fill_brewer(palette=col_pal)+
  scale_color_brewer(palette=col_pal) +
  #stat_bin(binwidth = 50000) +
  labs(title="Histogram for Price") +
  scale_x_continuous(name = "tax_amount", breaks = 10^5*(0:12), limits = c(0, 10^6)) +
  scale_y_continuous(name = "Frequency")

```

Linear buckets arn't cutting it
```{r fig.width=12, fig.height=8}
ggplot(df_properties, aes(tax_landvalue, tax_structurevalue))+ 
    geom_hex(bins = 50) + coord_fixed() +
    scale_fill_viridis() +
    theme_bw() +
  ggtitle("Frequency of Land vs. Structure Tax Value") +
  theme(legend.position="right") +
  xlim(0, 5*10^6) +
  ylim(0, 5*10^6)
  

```

Couldn't get viridis palette to work on a log scale hexbin so manually hacked my own using their color pallete.

Still not that informative, but we see that the vast majority of properties are taxed at <$100k for both land and property in 2015.

This is kinda a bummer, since I think the tax value may not gonnna be that close to the actual for sale value, wish I could buy houses in SoCal for <$100k.
```{r fig.width=12, fig.height=8}
ggplot(df_properties, aes(tax_landvalue, tax_structurevalue))+ 
    geom_hex(bins = 30) + coord_fixed() +
    scale_fill_gradientn(name = "Log Count", trans = "log10",
                        breaks = 10^(0:6), colors = c("#440154","#21908d","#fde725")) +
    theme_bw() +
  ggtitle("Log Frequency Land vs. Structure Tax Value") +
  theme(legend.position="right") +
  xlim(0, 5*10^6) +
  ylim(0, 5*10^6)
```

Using coordinate data we have a cool visual heatmap of where all our properties are in LA

```{r maphexbin, fig.width=12, fig.height=12}
df_properties[,lat:= latitude/1e06]
df_properties[,lon:= longitude/1e06]

ggplot(df_properties, aes(lat, lon))+ 
    geom_hex(bins = 200) + coord_fixed() +
    scale_fill_gradientn(name = "Number of Properties",
                        breaks = 250*(0:8), colors = c("#440154","#21908d","#fde725")) +
    theme_bw() +
  ggtitle("Log Frequency Land vs. Structure Tax Value") +
  theme(legend.position="right")

```


```{r map_data_prep}

lat <- range(df_properties$latitude/1e06,na.rm=T)
lon <- range(df_properties$longitude/1e06,na.rm=T)

tmp <- df_properties %>%
  select(id_parcel,longitude,latitude, tax_totalvalue) %>% 
  mutate(lon=longitude/1e6,lat=latitude/1e6, logtax = log10(tax_totalvalue)) %>% 
  select(id_parcel,lat,lon, logtax) %>% 
  left_join(df_transactions,by="id_parcel")

```

Learning the leaflet package wee!
This will help me make my interactive map of where I've traveled too.

To be continued
```{r include = FALSE, fig.width=12, fig.height=8}
#can't figure out colors yet
pal <- colorBin(viridis, tmp$logtax)


leaflet(tmp) %>% 
  addTiles() %>% 
  fitBounds(lon[1],lat[1],lon[2],lat[2]) %>% 
  addCircleMarkers(stroke=FALSE,
                   fillOpacity = 0.5,
                   clusterOptions = markerClusterOptions() 
                   #~pal(logtax)
                   ) %>% 
  addMiniMap()

```

```{r fig.width=12, fig.height=8}
library(ggmap)
mapa <- get_map(location = c(lon = -118.255, lat = 34.049), zoom = 9, maptype = "hybrid")    

ggmap(mapa) +
  geom_point(data = df_properties, aes(x = lon, y = lat, color = tax_totalvalue, alpha = 0.8))+
  scale_colour_gradient2(high = "darkred", mid = "yellow", low = "white", midpoint = median(df_properties$tax_totalvalue, na.rm = TRUE)/2, 
  space = "Lab", na.value = "grey50", guide = "colourbar") +
  scale_fill_viridis()

median(df_properties$tax_totalvalue, na.rm = TRUE)

hist(df_properties$tax_totalvalue)
```

